

```{r}
library(tidyverse)
library(here)
library(caret)
library(ggthemes)


```

```{r}
d <- read_csv(here("data/behavioral_data/trimmed_lt_d.csv"))
d <- d %>% 
  mutate(is_first_trial = trial_number == 1) %>% 
  mutate(violation_type_with_background = case_when(
    trial_type == "background" ~ "background", 
    TRUE ~ violation_type
  )) %>% 
  mutate(
    animacy = case_when(
    grepl("inanimate", background_type) ~ "inanimate", 
    TRUE ~ "animate"), 
    
    number = case_when(
    grepl("single", background_type) ~ "single", 
    TRUE ~ "pair"
  ),
  
    pose = case_when(
    grepl("left", background_type) ~ "left", 
    TRUE ~ "right"
  )
    )


```


# fix mixed effect model 

background as baseline

trial_number                                        -2.21e-02   3.73e-03  4.43e+04   -5.93  3.0e-09 ***
violation_type_with_backgroundpose                   1.70e-02   2.71e-02  4.43e+04    0.63  0.53159    
violation_type_with_backgroundanimacy                1.99e-01   2.80e-02  4.43e+04    7.12  1.1e-12 ***
violation_type_with_backgroundnumber                 3.25e-02   2.73e-02  4.43e+04    1.19  0.23327    
violation_type_with_backgroundidentity               2.11e-01   2.70e-02  4.43e+04    7.81  5.6e-15 ***
trial_number:violation_type_with_backgroundpose      1.48e-02   6.43e-03  4.43e+04    2.30  0.02124 *  
trial_number:violation_type_with_backgroundanimacy   2.27e-02   6.52e-03  4.43e+04    3.49  0.00049 ***
trial_number:violation_type_with_backgroundnumber    5.35e-03   6.44e-03  4.43e+04    0.83  0.40609    
trial_number:violation_type_with_backgroundidentity  1.01e-02   6.47e-03  4.43e+04    1.56  0.11901    

animacy as baseline

violation_type_with_backgroundpose                    -1.82e-01   3.74e-02  4.43e+04   -4.88  1.1e-06 ***
violation_type_with_backgroundbackground              -1.99e-01   2.80e-02  4.43e+04   -7.12  1.1e-12 ***
violation_type_with_backgroundnumber                  -1.67e-01   3.75e-02  4.43e+04   -4.45  8.7e-06 ***
violation_type_with_backgroundidentity                 1.17e-02   3.73e-02  4.43e+04    0.31  0.75485    
trial_number:violation_type_with_backgroundpose       -7.93e-03   8.61e-03  4.43e+04   -0.92  0.35725    
trial_number:violation_type_with_backgroundbackground -2.27e-02   6.52e-03  4.43e+04   -3.49  0.00049 ***
trial_number:violation_type_with_backgroundnumber     -1.74e-02   8.62e-03  4.43e+04   -2.02  0.04346 *  
trial_number:violation_type_with_backgroundidentity   -1.27e-02   8.64e-03  4.43e+04   -1.47  0.14270    

```{r}
d$violation_type_with_background = factor(d$violation_type_with_background, 
                                          levels = c("pose", 
                                                     "background",
                                                     "animacy", 
                                                     "number", 
                                                     "identity"))

m <- lmerTest::lmer(
  log(total_rt) ~ trial_number * is_first_trial + (1|prolific_id) + 
    trial_number * is_first_trial * number + 
    trial_number * is_first_trial * pose + 
    trial_number * is_first_trial * animacy + 
    trial_number * is_first_trial * violation_type_with_background + log(block_number),
  data = d 
)


(broom.mixed::tidy(m)) %>% saveRDS(here("cached_data/writing_cache/lmer_pose_baseline.Rds"))


```





